A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2021-11-07 paper/elsarticle/elsarticle-template.tex
paper

\begin{figure}
    \centering
    \subfloat[\centering Original image.]{{\includegraphics[height=3.4cm]{images/projim.png}}}
    \qquad
    \subfloat[\centering Projection to target location in scene plane (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
    \caption{Projection to a given target location in the scene plane (e.g., the floor, a wall, or table). (a) Original image. (b) Projecting the original image from the viewpoint of the projector (bottom left, black) to the scene plane introduces projective distortions if the projector is not fronto-parallel with the scene plane, as is the case here.}
    \label{fig:proj}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projection to target location in scene plane (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
    \caption{Projection to a given target location in the scene plane. (a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped with respect to the given target location using our approach. (b) Projecting the warped image from the viewpoint of the projector (bottom left, black) has the effect of projecting the \textit{original} image from the viewpoint of the virtual projector (center, gray). This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale. Note that black pixels will in effect be ignored by the projector.}
    \label{fig:warp}
\end{figure}

Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction. Notably, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face directly towards the scene plane (a camera we shall in what follows refer to as `downwards facing', for brevity). This facilitates placement of augmentations, since establishing the principal axes of the augmentation thus reduces to appropriately rotating the downwards-facing camera relative to the scene plane. We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}). We place this virtual projector to (i) face directly downwards to the target location in the scene plane and (ii) rotate its axes in accordance with those of the camera. Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane. Finally, our system is set up to handle a projector equipped with a steerable mirror in a manner that does not call for explicitly modeling the action of the steerable mirror on the projector, thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.

@@

Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro}---and exemplified in Figure~\ref{fig:proj}---can be achieved by means of an appropriate plane-induced homography \cite{Hartley2004}, computed by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view. This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene plane; modeling the geometry that relates projector and scene plane thus enables one to reason about the problem of correcting for distortions in terms of geometry. To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downwards to the scene plane and includes recovery of the scene plane as a convenient side effect. Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that warps the image to be projected in a manner that it appear undistorted to the viewer, placed in alignment with the axes of a proxy image of the scene, and at the desired metric scale. These two points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.

@@

@@

Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera then concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the \textit{projector} and the corresponding points in the scene. What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the downwards-facing camera (giving the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (giving the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$, for each projector calibration image).

\paragraph{2D-3D correspondences for camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane. The classical calibration surface is a chessboard pattern, and is such a pattern that we use as well. The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively. Note that the units of the chessboard's 3D points give the units of the camera calibration, and---since our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square. The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}. A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed. 2D-3D correspondences are then recovered for each calibration image, respectively, and the resulting list of correspondences is used to recover the camera calibration matrix~$\mathtt{K}_\text{cam}$.

\begin{figure}[ht]
    \centering
    \subfloat[\centering Projector calibration image (detections overlain).]{{\includegraphics[height=3cm]{images/detections.png} }}
    \qquad
    \subfloat[\centering 3D positions for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3cm]{images/2d3d.png}}}
    \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration. (a) For each target location, a projector calibration image is acquired from the viewpoint of the downwards-facing camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby. Circle pattern center points and chessboard corners detected automatically (overlain for illustration). (b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the camera (black) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown). The downwards-facing camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
    \label{fig:2d}
\end{figure}

\paragraph{2D-3D correspondences for projector calibration} For projector calibration, we obtain the 2D-3D correspondences by \textit{projecting} a calibration pattern to the scene plane, once for each target location. Since we know that the 3D positions of the projected pattern must lie in the scene plane, we can recover the 3D correspondence of a pattern point detected in the 2D image plane of the downwards-facing camera by intersecting the back-projection of the 2D position with the scene plane. We reuse the chessboard pattern---used above for camera calibration---to recover the scene plane locally to each target location via spatial resection \cite{Hartley2004}, by placing the chessboard pattern on the floor near the projected pattern and using the resulting 2D-3D chessboard correspondences as input to a PnP algorithm \cite{collins2014infinitesimal}. While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we recover a scene plane per projector calibration image to account for the possibility of an uneven floor. Since our projector calibration images thus already contain a chessboard, we choose an alternative calibration pattern as the pattern to project, namely an asymmetric circle pattern. We again use a specialized algorithm \cite{bradski2000opencv} to obtain the circle centers of the asymmetric circle pattern, capable of recovering circle centers even under projective distortions.

Given a projector calibration image acquired using the downwards-facing camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(a)); given the scene plane and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection
\begin{equation}
\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2
\end{equation}
of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:2d}(b)). We obtain the 2D positions of the 2D-3D correspondences by simply running the algorithm for detecting circle centers on the original asymmetric circles pattern image (i.e., the one we feed to the projector). Since---as in the case of recovering chessboard pattern corners---the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain respective 2D-3D correspondences for each projector calibration image, which we use to recover the projector calibration matrix~$\mathtt{K}_\text{proj}$.

Recall that the pose~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$ of the projector, for each projector calibration image~$j$, is provided alongside $\mathtt{K}_\text{proj}$, relative to the 3D points of the pattern and thus in the coordinate frame of the downwards-facing camera. Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror} (i.e., facing directly towards the target location). This is sufficient for our needs in Section~\ref{sec:approach:homography}, and it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_cam.pdf}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_cam.png}}}
    \caption{Obtaining the virtual camera. Our physical downwards-facing camera---in terms of which we wish to orient our augmentations---is almost certain to not face downwards precisely. Accordingly, we rotate the recovered downwards-facing camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal with respect to the minimum arc-length rotation. It is then, for each respective target position, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera. This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
    \label{fig:virtualcam}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_proj.pdf}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector. For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

\paragraph{Virtual projector} The placement of the virtual projector determines from which pose the image to be projected is to \textit{appear} to have been projected. We carry out this placement according to a small handful of steps. First, we rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual camera} facing directly\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}). It is to the virtual camera's $X$- and $Y$-axes that we subsequently set the $X$- and $Y$-axes of the virtual projector. To obtain the virtual projector, we (i) intersect the scene plane with the optical axis and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera. Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired projected image dimensions provided in metric units. The resulting virtual projector is thus rendered fronto-parallel with the scene plane---enabling projection to the scene plane absent of projective distortions---and projects to the scene plane with the desired scale.

It is, strictly speaking, in the sense that we set the $X$- and $Y$-axes of the virtual projector to the $X$- and $Y$-axes of the virtual camera that we place the horizontal and vertical axes of the desired augmentations in accordance with the horizontal and vertical image axes of the camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed. Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector. A steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, thereby facilitating placement.

\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location. Moreover, let $(\mathbf{n}^\top, -d)^\top \in \mathbb{P}^3$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X} \in \mathbb{R}^3$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$. The transformation that warps the image to be projected---by the recovered projector to the scene plane---such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
\begin{equation}
\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R} - \frac{\mathbf{t}\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
\label{homgen}
\end{equation}
a form of `plane-induced' 2D homography \cite{Hartley2004}. For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out intuitively with respect to the ground plane.

\begin{figure}
    \centering
    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=3.35cm]{images/system2.pdf}}}
    \quad
    \subfloat[\centering Projection of warped image from Figure~\ref{fig:warp}(a) to corresponding target location.]{{\includegraphics[height=3.35cm]{images/hans_undist.pdf}}}
    \caption{Evaluation scenario. (a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view. (b) Example augmentation produced by projecting---to one of our 15 target locations across the floorspace---an image warped using our approach, as seen from the downwards-facing camera. Note that the warp places the axes of the augmentation in accordance with the axes of the camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm). Corners of the full projected image extent projected to the floor in light green (emphasized with overlain dashed circles).}
    \label{fig:eval}
\end{figure}

\section{Evaluation}
We evaluate our approach by using a 960 pixel~$\times$~600 pixel image of the TU Wien logo (cf.\ Figure \ref{fig:warp}(a)) to augment 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria. We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace (i.e., the yellow and red lines in Figure~\ref{fig:2d}(b)), absent of projective distortions, and with the same metric dimensions of 50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}. All experiments were carried out by the same technician, experienced in both approaches.

>>>>

\begin{figure}
    \centering
    \subfloat[\centering Original image.]{{\includegraphics[height=3.4cm]{images/projim.png}}}
    \qquad
    \subfloat[\centering Projection to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table). (a) Original image, which we wish to project to the scene plane. (b) Projecting the original image to the scene plane from the viewpoint of the projector directed to the target location (bottom left, black with up vector red) introduces projective distortions if the projector is not fronto-parallel with the scene plane (as is the case here), with orientation and scale varying as a function of target location.}
    \label{fig:proj}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projection to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
    \caption{Projection to a given target location in the scene plane, correcting for projective distortions using our approach. (a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped using our approach with respect to the given target location. (b) Projecting the warped image to the scene plane from the viewpoint of the projector directed to the target location (bottom left, black with up vector red; same as in Figure~\ref{fig:proj}(b)) has the effect of projecting the \textit{original} image from the viewpoint of the \textit{virtual} projector (center, gray with up vector red). This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale. Note that black pixels will in effect be ignored by the projector.}
    \label{fig:warp}
\end{figure}

Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction. Notably, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face towards the scene plane, a camera we shall in what follows refer to as `downwards facing', for brevity. To compensate for the likely possibility that the camera not face downwards to the scene plane precisely, we rotate the recovered `downwards-facing' camera to face downwards directly, yielding what we refer to as the \textit{virtual} camera. Relying on the placement of the `downwards-facing' camera thus facilitates placement of augmentations, since setting the desired principal axes of augmentations with respect to the scene plane thus reduces to appropriately rotating the `downwards-facing' camera relative to the scene plane. We compute our corrective image warp for a given target location by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}). We place this virtual projector to (i) face directly downwards to the target location in the scene plane and (ii) rotate its axes to match those of the virtual camera. Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane. Finally, our system is set up for use with a projector equipped with a steerable mirror in a manner that does not call for explicitly modeling the action of the steerable mirror on the projector, thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.

@@

Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro}---and exemplified in Figure~\ref{fig:proj}---can be achieved by means of an appropriate plane-induced homography \cite{Hartley2004}, computed by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view. This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene plane; modeling the geometry that relates projector and scene plane thus enables one to reason about the problem of correcting for distortions in terms of geometry. To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downwards to the scene plane and includes recovery of the scene plane. Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that warps the image to be projected in a manner that it appear undistorted to the viewer, placed in alignment with the axes of a proxy image of the scene, and at the desired metric scale. These two points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.

@@

\begin{figure}[ht]
    \centering
    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[height=3.1cm]{images/detections.png} }}
    \qquad
    \subfloat[\centering Corresponding 3D positions for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.1cm]{images/2d3d.png}}}
    \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration. (a) For each target location, a projector calibration image is acquired from the viewpoint of the `downwards-facing' camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby. Circle pattern center points and chessboard corners detected automatically (overlain for illustration). (b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the camera (black with up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of the back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown). The `downwards-facing' camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
    \label{fig:2d}
\end{figure}

@@

Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera then concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the \textit{projector} and the corresponding points in the scene. What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the `downwards-facing' camera (giving the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (giving the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$, for each projector calibration image).

\paragraph{2D-3D correspondences for camera calibration} We recover 2D-3D correspondences in support of calibrating the `downwards-facing' camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane. The classical calibration surface is a chessboard pattern, and is such a pattern that we use as well. The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively. Note that the units of the chessboard's 3D points give the units of the camera calibration, and---since our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square. The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}. A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed. 2D-3D correspondences are then recovered for each calibration image, respectively, and the resulting list of correspondences is used to recover the camera calibration matrix~$\mathtt{K}_\text{cam}$.

\paragraph{2D-3D correspondences for projector calibration} For projector calibration, we obtain the 2D-3D correspondences by \textit{projecting} a calibration pattern to the scene plane, once for each target location. Since we know that the 3D positions of the projected pattern must lie in the scene plane, we can recover the 3D correspondence of a pattern point detected in the 2D image plane of the `downwards-facing' camera by intersecting the back-projection of the 2D position with the scene plane. We reuse the chessboard pattern---used above for camera calibration---to recover the scene plane locally to each target location via spatial resection \cite{Hartley2004}, by placing the chessboard pattern on the floor near the projected pattern and using the resulting 2D-3D chessboard correspondences as input to a PnP algorithm \cite{collins2014infinitesimal}. While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we recover a scene plane per projector calibration image to account for the possibility of an uneven floor. Since our projector calibration images thus already contain a chessboard, we choose an alternative calibration pattern as the pattern to project, namely an asymmetric circle pattern. We again use a specialized algorithm \cite{bradski2000opencv} to obtain the circle centers of the asymmetric circle pattern, capable of recovering circle centers even under projective distortions.

Given a projector calibration image acquired using the `downwards-facing' camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(a)); given the scene plane and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection
\begin{equation}
\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2
\end{equation}
of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:2d}(b)). We obtain the 2D positions of the 2D-3D correspondences by simply running the algorithm for detecting circle centers on the original asymmetric circles pattern image (i.e., the one we feed to the projector). Since---as in the case of recovering chessboard pattern corners---the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain respective 2D-3D correspondences for each projector calibration image, which we use to recover the projector calibration matrix~$\mathtt{K}_\text{proj}$.

Recall that the pose~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$ of the projector, for each projector calibration image~$j$, is provided alongside $\mathtt{K}_\text{proj}$, relative to the 3D points of the pattern and thus in the coordinate frame of the `downwards-facing' camera. Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror} (i.e., facing directly towards the target location). This is sufficient for our needs in Section~\ref{sec:approach:homography}, and it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_cam.pdf}}}
    \qquad
    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=4.2cm]{images/3_above3_cam.png}}}
    \caption{Obtaining the virtual camera. Our physical `downwards-facing' camera---in terms of whose rotation relative to the scene plane we wish to orient our augmentations---is almost certain to not face downwards precisely. Accordingly, we rotate the recovered `downwards-facing' camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation. It is then, for each respective target position, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera. This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
    \label{fig:virtualcam}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_proj.pdf}}}
    \qquad
    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector. For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera guaranteed---in contrast to the `downwards-facing' camera---to face \textit{directly} downwards to the scene plane (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

\paragraph{Virtual projector} The placement of the virtual projector determines from which pose the image to be projected is to \textit{appear} to have been projected.  We carry out this placement according to a small handful of steps. It is in accordance with the $X$- and $Y$-axes of the `downwards-facing' camera that we wish to place the $X$- and $Y$-axes of the augmentation; recognizing that a physical camera placed to face downwards is almost certain to not face downwards precisely, we first rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual} camera facing directly downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}). To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera. Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired projected image dimensions provided in metric units. These steps are illustrated in Figure~\ref{fig:virtualproj}. The resulting virtual projector is thus rendered fronto-parallel with the scene plane---enabling projection to the scene plane absent of projective distortions---and projects to the scene plane in the desired scale and with the the $X$- and $Y$-axes of the augmentation set in accordance with those of the `downwards-facing' camera. %\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, we align the virtual camera's optical axis exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.}

It is thus in the sense that we set the $X$- and $Y$-axes of the virtual projector to the $X$- and $Y$-axes of the virtual camera that we place the horizontal and vertical axes of the desired augmentations `in accordance' with the horizontal and vertical image axes of the `downwards-facing' camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed. Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector. A steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, thereby facilitating placement.

\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location. Moreover, let $(\mathbf{n}^\top, -d)^\top \in \mathbb{P}^3$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X} \in \mathbb{R}^3$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$. The transformation that warps the image to be projected---by the recovered projector to the scene plane---such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}) is given the by the $3~\times~3$ matrix
\begin{equation}
\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R} - \frac{\mathbf{t}\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
\label{homgen}
\end{equation}
a form of `plane-induced' 2D homography \cite{Hartley2004}. For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out intuitively with respect to the ground plane.

\begin{figure}
    \centering
    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=3.35cm]{images/system2.pdf}}}
    \quad
    \subfloat[\centering Projection of warped image from Figure~\ref{fig:warp}(a) to corresponding target location.]{{\includegraphics[height=3.35cm]{images/hans_undist.pdf}}}
    \caption{Evaluation scenario. (a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view. (b) Example augmentation produced by projecting---to one of our 15 target locations across the floorspace---an image warped using our approach, as seen from the `downwards-facing' camera. Note that the warp places the axes of the augmentation in accordance with the axes of the camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm). Corners of the full projected image extent projected to the floor in light green (emphasized with overlain dashed circles).}
    \label{fig:eval}
\end{figure}

\section{Evaluation}
We evaluate our approach by using a 960 pixel~$\times$~600 pixel image of the TU Wien logo (cf.\ Figure \ref{fig:proj}(a)) to augment 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria. We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image of the TU Wien logo for each of the 15 locations, aligned with the principal axes of the floorspace (i.e., the yellow and red lines in Figure~\ref{fig:2d}(b)), absent of projective distortions, and with the same metric dimensions of 50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}. All experiments were carried out by the same technician, experienced in both approaches.
